---
title: "Craft Beers"
author: "Linda Eliasen/Limin Zheng"
date: "February 23, 2019"
output: html_document
---

```{r setup, include=FALSE}
# Setting our Echo on output for the presentation to be false
knitr::opts_chunk$set(echo = FALSE)
```


```{r libraries, include=FALSE}
library(cowplot)
library(dplyr)
library(tidyr)
library(knitr)
library(readr)
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(stringr)
library(kableExtra)
library(highcharter)
library(dataMaid)
```

## bringing in data files
```{r files}

brew<- read.csv("breweries.csv" , header=TRUE, stringsAsFactors=FALSE)
  colnames(brew)<- c('Brewery_id', 'Name', 'City', 'State')

beer <- read.csv("beers.csv" , header=TRUE, stringsAsFactors=FALSE)
  
brew<-merge(brew,beer, by="Brewery_id",all.x = TRUE)
colnames(brew)<- c('Brewery_id', 'Brewery', 'City', 'State', 'Beer', 'Beer_ID', 'ABV', 'IBU', 'Style', 'Ounces')
brew$IBU <- as.numeric(brew$IBU)

sbystate <- read.csv("StateStats.csv" , header=TRUE, stringsAsFactors=FALSE)
sbystate$Economic.Impact.Mil<-as.numeric(sbystate$Economic.Impact.Mil)
sbystate$Economic.Impact.Rank<-as.numeric(sbystate$Economic.Impact.Rank)
sbystate$Barrels.Per.Year<-as.numeric(sbystate$Barrels.Per.Year)
```


Top and Bottom of Brewery Data Set

```{r head}
brew <- brew %>% 
	select(Brewery, Beer, Style, ABV, IBU, Ounces, City, State, Beer_ID, Brewery_id)

kable(head(brew,6))%>%
  kable_styling(full_width = F)
```

```{r tail}
kable(tail(brew,6))%>%
  kable_styling(full_width = F)
```


Report the number of NA's in each column.
```{r missing}
sha<-colSums(is.na(brew))
kable(sha)%>%
  kable_styling(full_width = F)
```


How many breweries are present in each state? Compute the median alcohol content and international bitterness unit for each state, plus adding in other data elements from third file
```{r facts}
state <-as.data.frame(brew%>% group_by(State) 
  %>% summarise(MedianABV=median(ABV, na.rm=TRUE), 
  MedianIBU=median(IBU, na.rm=TRUE),
  Breweries = n_distinct(Brewery_id)))

state$Breweries<- as.numeric(state$Breweries)

state<-merge(state, sbystate, by="State",all.x = TRUE)
colnames(state)<- c('State', 'MedianABV', 'MedianIBU', 'Breweries','BreweriesPerCapita', 'EconomicImpactM$', 'EconomicImpactRank', 'BarrelsPerYear', 'GallonsPerAdult')

kable(state)%>%
  kable_styling(full_width = F)
```


Heat map number of breweries by state
```{r mapBreweries}

mapdata <- get_data_from_map(download_map_data("countries/us/us-all"))

#need to remove extra spaces from state abreviation
state$State <- str_trim(state$State,side="left")
state <- state %>% mutate(`hc-a2`= State);

hcmap("countries/us/us-all", data = state, value = "Breweries",
	  joinBy = "hc-a2", name = "Breweries",
	  dataLabels = list(enabled = TRUE, format = '{point.name}'),
	  borderColor = "#FAFAFA", borderWidth = 0.1,
	  tooltip = list(valueDecimals = 0,valueSuffix = " Breweries")) %>%
	hc_mapNavigation(enabled = TRUE) %>%
	hc_title(text = "<b>Number of Breweries by State</b>",
			 margin = 20, align = "center", 
			 style = list(color = "#000000", 
			 			 useHTML = TRUE))
```

Plot a bar chart to compare.

Median Alcohol content
```{r abv}
#Graph ABV
ggplot(state, aes(x=reorder(State, -MedianABV), y=MedianABV, fill=MedianABV)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Median ABV") +
  ggtitle("Median ABV by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
coord_flip() 
```

Median International Bitterness Level by State
```{r ibu}
 #Graph IBU
ggplot(state, aes(x=reorder(State, -MedianIBU), y=MedianIBU, fill=MedianIBU)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Median IBU") +
  ggtitle("Median IBU by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
coord_flip()
```


charts side by side... needs widening
```{r abv_ibu}
#Graph side by side
abv<-ggplot(state, aes(x=reorder(State, -MedianABV), y=MedianABV, fill=MedianABV)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Median ABV") +
  ggtitle("Median ABV by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  coord_flip() 

ibu<-ggplot(state, aes(x=reorder(State, -MedianIBU), y=MedianIBU, fill=MedianIBU)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Median IBU") +
  ggtitle("Median IBU by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  coord_flip()

plot_grid(abv, ibu,labels = "AUTO")
```

Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r max}
ma <- data.frame(brew[order(-brew$ABV),])
ma=ma[1,c(7,2,4,5)]
ma<-ma[1, ]
mi <- data.frame(brew[order(-brew$IBU),])
mi=mi[1,c(7,2,4,5)]
mi<-mi[1, ]

top<-rbind(ma, mi)
top[is.na(top)] <- ""

kable(head(top[,1:4]),
      format = "pandoc", 
      digits = 3,        
      caption = "State with Highest ABV and Most Bitter Beer",
      col.names = c("State","Beer","ABV","IBV"),
      row.names = FALSE,
      align = c("l"))  
```

Summary statistics for the ABV variable (also adding in IBU)
```{r summary}
AlcoholbyVolume<-summary(brew$ABV, na.rm=TRUE, digits = 3)
InternationalBitterness<-summary(brew$IBU, na.rm=TRUE)

EconomicImpactInMillions<-summary(state$`EconomicImpactM$`)#, digits=1)
AnnualProductionBarrels<-summary(state$BarrelsPerYear)#, digits=0)

avib<-rbind(AlcoholbyVolume, InternationalBitterness)
     
kable(avib)%>%
  kable_styling(full_width = F)

eco<-rbind(EconomicImpactInMillions, AnnualProductionBarrels)
     
kable(eco)%>%
  kable_styling(full_width = F)
```

Is there an apparent relationship between the bitterness of the beer and its alcoholic content? 
```{r scatter}
ggplot(brew, aes(IBU, ABV)) +
geom_point(color="blue") +
xlim(0,150)+ylim(0,.150)+
  labs(x="IBU",y="ABV") +
ggtitle("Bitterness v. Alcohol Content") +
geom_smooth(method=lm, color="black", se=FALSE) 
```

ADDITIONAL CHARTS

Number of different craft beers in each state
```{r diffbeer}
btype<-as.data.frame(brew %>% group_by(State)%>% summarise(BeerTypes = n_distinct(Beer)))
btype<-btype[order(-btype$BeerTypes),]
kable(btype)%>%
  kable_styling(full_width = F)

ggplot(btype, aes(x=reorder(State, -BeerTypes), y=BeerTypes, fill=BeerTypes)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Number") +
  ggtitle("Number of Different Craft Beers by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  coord_flip()

```


Number of different beer STYLES in each state
```{r style}

bstyl<-as.data.frame(brew %>% group_by(State)%>% summarise(BeerStyles = n_distinct(Style)))
bstyl<-bstyl[order(-bstyl$BeerStyles),]
kable(bstyl)%>%
  kable_styling(full_width = F)

ggplot(bstyl, aes(x=reorder(State, -BeerStyles), y=BeerStyles, fill=BeerStyles)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Number") +
  ggtitle("Number of Different Beer Styles by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  coord_flip()
```

Beer and Brand Style by State
```{r brndstyl}
BrndStyl <- merge(btype, bstyl, by='State')
BrndStyl<-BrndStyl[order(-BrndStyl$BeerStyles),]

kable(BrndStyl)%>%
  kable_styling(full_width = F)

```


Treemap of Beer Styles
```{r tmapstyle}
brew%>%
  count(Style)%>%
  arrange(n)%>%
  hchart(type = "treemap", hcaes(x = Style, value = n, color = n))
```


```{r }
statesum<-merge(state, BrndStyl, by="State",all.x = TRUE)
colnames(statesum)<- c('State', 'MedianABV', 'MedianIBU', 'Breweries','BreweriesPerCapita', 'EconomicImpactMil', 'EconomicImpactRank', 'BarrelsPerYear', 'GallonsPerAdult', 'DifferentTypes', 'DifferentStyles')

statesum<-statesum[,c(1,2,3,4,10,11,5,6,7,8,9)]

#kable(statesum)%>%
#  kable_styling(full_width = F)

```

State Top 10 Rank by Economic Impact
```{r rank}
rankei<-statesum[order(-statesum$EconomicImpactRank),]
rankei=rankei[,c(1,9)]

kable(head(rankei,10))%>%
  kable_styling(full_width = F)

```

Revenue by State in M$
```{r rev}
rev<-statesum[order(-statesum$EconomicImpactMil),]
rev=rev[,c(1,8)]

kable(head(rev,10))%>%
  kable_styling(full_width = F)

ggplot(rev, aes(x=reorder(State, -EconomicImpactMil), y=EconomicImpactMil, fill=EconomicImpactMil)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Revenue in M$") +
  ggtitle("Craft Beer Revenue M$ by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  coord_flip()

```


Gallons Per Adult 21+ by State
```{r gallon}
cons<-statesum[order(-statesum$GallonsPerAdult),]
cons=cons[,c(1,11)]

kable(head(cons,10))%>%
  kable_styling(full_width = F)

ggplot(cons, aes(x=reorder(State, -GallonsPerAdult), y=GallonsPerAdult, fill=GallonsPerAdult)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Gallons") +
  ggtitle("Gallons Produced Per Adult by State") +
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  coord_flip()

```

write.csv(statesum,'statesum.csv')

#####Sources
*https://www.datacamp.com/community/tutorials/data-visualization-highcharter-r*
