---
title: "Beers & Breweries"
author: "Linda Eliasen/Limin Zheng"
date: "February 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(ggpubr)
library(RColorBrewer)
library(bindrcpp)
library(cowplot)
library(kableExtra)
```

```{r}
#BREWERIES Import
brew<- read.csv("breweries.csv" , header=TRUE)
```

How many breweries are present in each state?
```{r}
bbs <-(brew%>% group_by(State) %>% count(State))
bbs<-as.data.frame(bbs)

kable(bbs)%>%
  kable_styling(full_width = F)
```

Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
```{r}
df<- read.csv("beers.csv" , header=TRUE)
df <- df[order(df$Beer_ID),]
```

```{r}
brew<-merge(brew,df, by.x="Brew_ID", by.y="Brewery_id",all.x = TRUE)
colnames(brew)<- c('Brewery_id', 'Name', 'City', 'State', 'Beer', 'Beer_ID', 'ABV', 'IBU', 'Style', 'Ounces')

brew$ABV <- as.numeric(brew$ABV)
brew$IBU <- as.numeric(brew$IBU)

```

```{r}
#First 6 observations
kable(head(brew,6))%>%
  kable_styling()

#Last 6 observations
kable(tail(brew,6))%>%
  kable_styling()
```

Report the number of NA's in each column.
```{r}
sha<-colSums(is.na(brew))
kable(sha)%>%
  kable_styling(full_width = F)
```

Compute the median alcohol content and international bitterness unit for each state. 
```{r}
beers <-as.data.frame(brew%>% group_by(State) %>% summarise(MedianABV=median(ABV, na.rm=TRUE), 
  MedianIBU=median(IBU, na.rm=TRUE)))

kable(beers)%>%
  kable_styling(full_width = F)
```

Plot a bar chart to compare.

Median Alcohol content
```{r}
#Graph ABV
ggplot(beers, aes(x=reorder(State, -MedianABV), y=MedianABV, fill=MedianABV)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Median ABV") +
  ggtitle("Median ABV by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
coord_flip() 
```

Median International Bitterness Level by State
```{r}
 #Graph IBU
ggplot(beers, aes(x=reorder(State, -MedianIBU), y=MedianIBU, fill=MedianIBU)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Median IBU") +
  ggtitle("Median IBU by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
coord_flip()
```

Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
ma <- data.frame(brew[order(-brew$ABV),])
ma=ma[1,c(4,5,7,8)]
ma[1, ]
mi <- data.frame(brew[order(-brew$IBU),])
mi=mi[1,c(4,5,7,8)]
mi[1, ]
top<-rbind(ma, mi)
top[is.na(top)] <- ""

kable(head(top[,1:4]),
      format = "pandoc", 
      digits = 3,        
      caption = "State with Highest ABV and Most Bitter Beer",
      col.names = c("State","Beer","ABV","IBV"),
      row.names = FALSE,
      align = c("l"))  
```

Summary statistics for the ABV variable.
```{r}
summary(brew$ABV, na.rm=TRUE)
```


Is there an apparent relationship between the bitterness of the beer and its alcoholic content? 
```{r}
ggplot(brew, aes(IBU, ABV)) +
geom_point(color="blue") +
xlim(0,150)+ylim(0,.150)+
  labs(x="IBU",y="ABV") +
ggtitle("Bitterness v. Alcohol Content") +
geom_smooth(method=lm, color="black", se=FALSE) 
```



ADDITIONAL CHARTS

Number of different craft beers in each state
```{r}
btype<-as.data.frame(brew %>% group_by(State)%>% summarise(BeerTypes = n_distinct(Beer)))
btype<-btype[order(-btype$BeerTypes),]
kable(btype)%>%
  kable_styling(full_width = F)

ggplot(btype, aes(x=reorder(State, -BeerTypes), y=BeerTypes, fill=BeerTypes)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Number") +
  ggtitle("Number of Different Craft Beers by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  coord_flip()

```

Number of different beer STYLES in each state
```{r}

bstyl<-as.data.frame(brew %>% group_by(State)%>% summarise(BeerStyles = n_distinct(Style)))
bstyl<-bstyl[order(-bstyl$BeerStyles),]
kable(bstyl)%>%
  kable_styling(full_width = F)

ggplot(bstyl, aes(x=reorder(State, -BeerStyles), y=BeerStyles, fill=BeerStyles)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Number") +
  ggtitle("Number of Different Beer Styles by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  coord_flip()
```


Beer and Brand Style by State
```{r}
BrndStyl <- merge(btype, bstyl, by='State')
BrndStyl<-BrndStyl[order(-BrndStyl$BeerStyles),]
kable(BrndStyl)%>%
  kable_styling(full_width = F)

```

