---
title: "Craft Beer Market Analysis"
author: "Team Members: Linda Eliasen and Limin Zheng"
date: "February 2019"
output: html_document
---

```{r setup, include=FALSE}
# Setting our Echo on output for the presentation to be false
knitr::opts_chunk$set(echo = FALSE, messages = FALSE, warning = FALSE)
```


```{r libraries, include=FALSE}
library(cowplot)
library(dplyr)
library(tidyr)
library(knitr)
library(readr)
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(stringr)
library(kableExtra)
library(highcharter)
library(dataMaid)
```

### Data Overview
```{r files}

brew<- read.csv("breweries.csv" , header=TRUE, stringsAsFactors=FALSE)
  colnames(brew)<- c('Brewery_id', 'Name', 'City', 'State')

beer <- read.csv("beers.csv" , header=TRUE, stringsAsFactors=FALSE)
  
brew<-merge(brew,beer, by="Brewery_id",all.x = TRUE)
colnames(brew)<- c('Brewery_id', 'Brewery', 'City', 'State', 'Beer', 'Beer_ID', 'ABV', 'IBU', 'Style', 'Ounces')
brew$IBU <- as.numeric(brew$IBU)

sbystate <- read.csv("StateStats.csv" , header=TRUE, stringsAsFactors=FALSE)
sbystate$Economic.Impact.Mil<-as.numeric(sbystate$Economic.Impact.Mil)
sbystate$Economic.Impact.Rank<-as.numeric(sbystate$Economic.Impact.Rank)
sbystate$Barrels.Per.Year<-as.numeric(sbystate$Barrels.Per.Year)
```


Top and Bottom of Brewery Data Set

```{r head}
brew <- brew %>% 
	select(Brewery, Beer, Style, ABV, IBU, Ounces, City, State, Beer_ID, Brewery_id)

kable(head(brew,6))%>%
  kable_styling(full_width = F)
```

```{r tail}
kable(tail(brew,6))%>%
  kable_styling(full_width = F)
```


Report the number of NA's in each column.
```{r missing}
sha<-colSums(is.na(brew))
kable(sha)%>%
  kable_styling(full_width = F)
```

CRAFT BEER DEMOGRAPHICS BY STATE
How many breweries are present in each state? Compute the median alcohol content and international bitterness unit for each state, plus adding in other data elements from third file
```{r facts}
state <-as.data.frame(brew%>% group_by(State) 
  %>% summarise(MedianABV=median(ABV, na.rm=TRUE), 
  MedianIBU=median(IBU, na.rm=TRUE),
  Breweries = n_distinct(Brewery_id)))

state$Breweries<- as.numeric(state$Breweries)

state<-merge(state, sbystate, by="State",all.x = TRUE)
colnames(state)<- c('State', 'MedianABV', 'MedianIBU', 'Breweries','BreweriesPerCapita', 'EconomicImpactM$', 'EconomicImpactRank', 'BarrelsPerYear', 'GallonsPerAdult')

kable(state)%>%
kable_styling(bootstrap_options = "striped", "condensed")
```


Heat map number of breweries by state
```{r mapBreweries}

mapdata <- get_data_from_map(download_map_data("countries/us/us-all"))

#need to remove extra spaces from state abreviation
state$State <- str_trim(state$State,side="left")
state <- state %>% mutate(`hc-a2`= State);

hcmap("countries/us/us-all", data = state, value = "Breweries",
	  joinBy = "hc-a2", name = "Breweries",
	  dataLabels = list(enabled = TRUE, format = '{point.name}'),
	  borderColor = "#FAFAFA", borderWidth = 0.1,
	  tooltip = list(valueDecimals = 0,valueSuffix = " Breweries")) %>%
	hc_mapNavigation(enabled = TRUE) %>%
	hc_title(text = "<b>Number of Breweries by State</b>",
			 margin = 20, align = "center", 
			 style = list(color = "#000000", 
			 			 useHTML = TRUE))
```

Heat map number of breweries per capita by state
```{r mapBrewPC}

mapdata <- get_data_from_map(download_map_data("countries/us/us-all"))

#need to remove extra spaces from state abreviation
state$State <- str_trim(state$State,side="left")
state <- state %>% mutate(`hc-a2`= State);

hcmap("countries/us/us-all", data = state, value = "BreweriesPerCapita",
	  joinBy = "hc-a2", name = "BreweriesPerCapita",
	  dataLabels = list(enabled = TRUE, format = '{point.name}'),
	  borderColor = "#FAFAFA", borderWidth = 0.1,
	  tooltip = list(valueDecimals = 0,valueSuffix = " Breweries")) %>%
	hc_mapNavigation(enabled = TRUE) %>%
	hc_title(text = "<b>Breweries Per Capita<b>",
			 margin = 20, align = "center", 
			 style = list(color = "#000000", 
			 			 useHTML = TRUE))
```




Plot a bar chart to compare Median ABV and IBU by State
```{r }
ggplot(state, aes(x=reorder(State, -MedianABV),y=MedianABV, fill=MedianIBU))+labs(title="Plot of ABV by state", x="State", y="MedianABV", fill="median IBU")+geom_bar(stat="identity") + theme_bw() +theme(axis.text.x=element_text(angle=90, size = 7)) +theme(plot.title = element_text(hjust = 0.5))+theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+coord_flip()
```

charts side by side... still work in progress, want to turn into back to back
```{r abv_ibu}
#Graph side by side
abv<-ggplot(state, aes(x=State, y=MedianABV, fill=MedianABV)) + 
  geom_bar(stat="identity") +
  ggtitle("Median ABV by State") +
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=4.5))+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(colour="black", size =6))+
  theme(legend.title = element_blank()) +
  coord_flip() 

ibu<-ggplot(state, aes(x=State, y=MedianIBU, fill=MedianIBU)) + 
  geom_bar(stat="identity") +
  ggtitle("Median IBU by State") +
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=4.5))+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(colour="black", size =8))+
  theme(legend.title = element_blank()) +
  coord_flip()

plotf<-plot_grid(abv, ibu, align="hv", nrow = 1, ncol = 2) 

plotf
```

SUMMARY DATA
Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r max}
ma <- data.frame(brew[order(-brew$ABV),])
ma=ma[1,c(8,2,4,5)]
ma<-ma[1, ]
mi <- data.frame(brew[order(-brew$IBU),])
mi=mi[1,c(8,2,4,5)]
mi<-mi[1, ]

top<-rbind(ma, mi)
top[is.na(top)] <- ""

kable(head(top[,1:4]),
      format = "pandoc", 
      digits = 3,        
      caption = "State with Highest ABV and Most Bitter Beer",
      col.names = c("State","Beer","ABV","IBU"),
      row.names = FALSE,
      align = c("l"))  
```

Summary statistics for key variables
```{r summary}
Alcohol_by_Volume<-(summary(brew$ABV, na.rm=TRUE, digits=3 ))

International_Bitterness_Units<-(summary(brew$IBU, na.rm=TRUE))

abv<-rbind(Alcohol_by_Volume)
kable(abv)%>%
 kable_styling(full_width = F)

ibu<-rbind(International_Bitterness_Units)
kable(ibu)%>%
  kable_styling(full_width = F)

```

Histogram with density plot ABV
```{r abvhist}
ggplot(brew, aes(x=ABV)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="lightblue", 
                position="identity")+
 	geom_density(alpha=.2, color="blue")+
	geom_vline(aes(xintercept=mean(brew$ABV, na.rm = T)), color="red",
             linetype="longdash")+
	labs(title="Alcohol by Volume",x="ABV", y = "Density")+
	geom_text(data=brew, mapping=aes(x=mean(brew$ABV, na.rm = T), y=50, label=round(mean(brew$ABV, na.rm = T),digits=4)), size=4, hjust=-0)
```


Histogram with density plot IBU
```{r ibuhist}
ggplot(brew, aes(x=IBU)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="lightblue", 
                position="identity")+
 	geom_density(alpha=.2, color="blue")+
	geom_vline(aes(xintercept=mean(brew$IBU, na.rm = T)), color="red",
             linetype="longdash")+
	labs(title="International Bitterness Unit",x="IBU", y = "Density")+
	geom_text(data=brew, mapping=aes(x=round(mean(brew$IBU, na.rm = T),digits=2), y=0.022, label=round(mean(brew$IBU, na.rm = T),digits=1)), size=4, hjust=0)
```


#TEST GRAPH FOR REGRESSION

```{r}
ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point(color="blue") +
xlim(0,150)+ylim(0,.150)+
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}

ggplotRegression(lm(ABV~IBU, data = brew))
```




Is there an apparent relationship between the bitterness of the beer and its alcoholic content? 
```{r scatter}
ggplot(brew, aes(IBU, ABV)) +
geom_point(color="blue") +
xlim(0,150)+ylim(0,.150)+
  labs(x="IBU",y="ABV") +
ggtitle("Bitterness v. Alcohol Content") +
geom_smooth(method=lm, color="black", se=FALSE)

```

The p-value of the test is less than the significance level alpha = 0.05. We can conclude that ABV and IBU are correlated with a correlation coefficient of 0.67.
```{r corpear}
res <- cor.test(brew$IBU, brew$ABV, 
                    method = "pearson", use = "complete.obs")
res
#cor(brew, x=IBU, y=AVB,  method = "pearson", use = "complete.obs")
```



Treemap of Beer Styles
```{r tmapstyle}
brew%>%
  count(Style)%>%
  arrange(n)%>%
  hchart(type = "treemap", hcaes(x = Style, value = n, color = n))
```

Top 10 Craft Beer Styles
```{r stlexp}

bstlexp<-as.data.frame(brew %>% group_by(Style)
  %>% summarise(Median_ABV=median(ABV, na.rm=TRUE), 
 Median_IBU=median(IBU, na.rm=TRUE),
  Beers = n_distinct(Beer)))

bstlexp<-bstlexp[order(-bstlexp$Beers),]

bstlexp=bstlexp[,c(1,4,2,3)]

#kable(bstlexp)%>%
 # kable_styling(full_width = F)

#Top 10 Styles in the United States
kable(head(bstlexp,10))%>%
  kable_styling(full_width = F)

```


Number of different beer STYLES in each state
```{r style}

bstyl<-as.data.frame(brew %>% group_by(State)%>% summarise(BeerStyles=n_distinct(Style)))
bstyl<-bstyl[order(-bstyl$BeerStyles),]

bst<-ggplot(bstyl, aes(x=State, y=BeerStyles, fill=BeerStyles)) + 
  geom_bar(stat="identity") +
  ggtitle("Craft Beer Styles") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(colour="black", size =8))+
  theme(legend.title = element_blank()) +
  coord_flip()

btype<-as.data.frame(brew %>% group_by(State)
        %>% summarise(BeerTypes = n_distinct(Beer)))

btp<-ggplot(btype, aes(x=State, y=BeerTypes, fill=BeerTypes)) + 
  geom_bar(stat="identity") +
  ggtitle("Craft Beers") +
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(colour="black", size =8))+
  theme(legend.title = element_blank()) +
  coord_flip()

plotb<-plot_grid(bst, btp, align="hv", nrow = 1, ncol = 2) 

plotb
```

Heat map Revenue in Mill$ by state
```{r revmap}
BrndStyl <- merge(btype, bstyl, by='State')

statesum<-merge(state, BrndStyl, by="State",all.x = TRUE)
colnames(statesum)<- c('State', 'MedianABV', 'MedianIBU', 'Breweries','BreweriesPerCapita', 'EconomicImpactMil', 'EconomicImpactRank', 'BarrelsPerYear', 'GallonsPerAdult', 'DifferentTypes', 'DifferentStyles')
statesum<-statesum[,c(1,2,3,4,10,11,5,6,7,8,9)]

reva<-statesum[order(-statesum$EconomicImpactMil),]
reva=reva[,c(1,8)]


mapdata <- get_data_from_map(download_map_data("countries/us/us-all"))

reva$State <- str_trim(reva$State,side="left")
rev <- reva %>% mutate(`hc-a2`= State);


hcmap("countries/us/us-all", data = rev, value = "EconomicImpactMil",
	  joinBy = "hc-a2", name = "EconomicImpactMil",
	  dataLabels = list(enabled = TRUE, format = '{point.name}'),
	  borderColor = "#FAFAFA", borderWidth = 0.1,
	  tooltip = list(valueDecimals = 0,valueSuffix = "Economic Impact")) %>%
	hc_mapNavigation(enabled = TRUE) %>%
	hc_title(text = "<b>Economic Impact<b>",
			 margin = 20, align = "center", 
			 style = list(color = "#000000", 
			 			 useHTML = TRUE))
```
Summary statistics for key variables
```{r summary2}

Economic_Impact_In_Millions<-summary(state$`EconomicImpactM$`)

Barrels_Produced_Annually<-summary(state$BarrelsPerYear)

eco<-rbind(Economic_Impact_In_Millions)
kable(eco)%>%
  kable_styling(full_width = F)

prd<-rbind(Barrels_Produced_Annually)
kable(prd)%>%
  kable_styling(full_width = F)
```


Top 10 Revenue by State in M$
```{r rev}

kable(head(reva,10))%>%
  kable_styling(full_width = F)
```

Top 10 States - Barrel Production
```{r barrels}
bars<-statesum[order(-statesum$BarrelsPerYear),]
bars=bars[,c(1,10)]

kable(head(bars,10))%>%
  kable_styling(full_width = F)

```




APPENDIX

Craft Beer Styles in order of popularity
```{r stlexp2}

kable(bstlexp)%>%
kable_styling(full_width = F)
```


Revenue by State in M$
```{r revF}

ggplot(reva, aes(x=reorder(State, -EconomicImpactMil), y=EconomicImpactMil, fill=EconomicImpactMil)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Revenue in M$") +
  ggtitle("Craft Beer Revenue M$ by State") +
  theme(plot.title = element_text(hjust = 0.5))+ #forces centering
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  coord_flip()

kable(reva)%>%
  kable_styling(full_width = F)
```


Beer and Brand Style by State
```{r brndstyl2}
kable(BrndStyl)%>%
  kable_styling(full_width = F)

```


Gallons Per Adult 21+ by State
```{r gallon}
cons<-statesum[order(-statesum$GallonsPerAdult),]
cons=cons[,c(1,11)]


ggplot(cons, aes(x=reorder(State, -GallonsPerAdult), y=GallonsPerAdult, fill=GallonsPerAdult)) + 
  geom_bar(stat="identity") +
  labs(x="State",y="Gallons") +
  ggtitle("Gallons Produced Per Adult by State") +
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme(axis.text.y = element_text(hjust=1, vjust=0.5, size=5))+
  coord_flip()


kable(cons)%>%
  kable_styling(full_width = F)

```

Revenue vs. breweries per state
```{r scatter2 }
ggplot(state, aes(Breweries, `EconomicImpactM$`)) +
geom_point(color="blue") +
xlim(0,50)+ylim(0,8000)+
  labs(x="Breweries",y="EconomicImpactM$") +
ggtitle("EconomicImpactM$ v. Breweries") +
geom_smooth(method=lm, color="black", se=FALSE) +
geom_text(aes(label=State),hjust=0, vjust=0, size=3)

```


 breweries per capita vs. Gallon per adult in each state
```{r scatter3}
ggplot(state, aes(GallonsPerAdult, BreweriesPerCapita)) +
geom_point(color="blue") +
xlim(0,15)+ylim(0,10)+
  labs(x="GallonsPerAdult",y="BreweriesPerCapita") +
ggtitle("Breweries Per Capita v. Gallons Per Adult") +
geom_smooth(method=lm, color="black", se=FALSE)+ 
  geom_text(aes(label=State),hjust=0, vjust=0, size=3)

```



#####Sources
*https://www.brewersassociation.org/*
*https://www.highcharts.com/docs/maps/map-collection*